---
title: "NSF Graduate Student Data Exploration"
author: "Qi Wang"
date: "2023-09-18"
output: pdf_document
---

```{r setup, include=FALSE}
rm(list = ls())
knitr::opts_chunk$set(echo = TRUE)
library(openxlsx)
library(dplyr)
library(nlme)
library(ggplot2)
```

# Data introduction

In each file, there are three sheets, recording the race, support and postdoctoral. Each sheet contains a unique combination of year, school_id, and gss_code. These ID information are replicated in each sheet in each file, meaning that we are researching the same bunch of schools, the only difference is the variables. 

There is a version of 2017 called "2017old" since in 2017 they adjusted some criteria.Due to these changes, the 2017 data are not comparable to prior years. A set of bridge estimates was created to permit comparisons to previous years and for trend analyses.

To begin with, I will first go further into the first sheet of each file and make it as an example about data wrangling for further reference.

```{r}
get_data <- function(var_name_old, var_name_new, sheet){
  # Since there is a data collection method change in 2017, so var_name_old should be the previous variable name, and the var_name_new should be the name after 2017. There is a sheet recording the equivalent names for them.
  ID_variables <- c(
    "institution_id","school_id","UNITID","year", "Institution_Name","hdg_inst","toc_code", "institution_state", "hbcu_flag","land_grant_flag","carnegie_code_1994","carnegie_code_2005","carnegie_code_2010","carnegie_code_2015","full_school_name","school_name","school_zip","school_type_code","gss_code","hdg_code"
  )
  
  all_sheet<- vector("list")
  selected_mat <-  vector("list")
for (i in 1:50) {
  print( paste("Now retrieving data of year: ",i+1971))
    all_sheet[[i]] <- read.xlsx( paste("D:/77/UCSC/study/Reasearch/temp/NSF_dat/NSF_grad/gss",1971+i,"_Code.xlsx", sep = ""), sheet = sheet)
    if(1971+i <= 2016){selected_mat[[i]] <- all_sheet[[i]][,c(ID_variables, var_name_old)]}
    else{selected_mat[[i]] <- all_sheet[[i]][,c(ID_variables, var_name_new)]}
}
  final_mat <- bind_rows(selected_mat)
  return(final_mat)
}


dat <- get_data("ft_tot_all_races_v", "ft_tot_all_races_v", "Race")

```

```{r}
use_dat <- dat[-which(dat$UNITID==999999),c("school_id","year","gss_code","ft_tot_all_races_v")]

# summed_dat <- aggregate( ft_tot_all_races_v ~  school_id + gss_code + year, data = use_dat, FUN = function(x) sum(x, na.rm = TRUE))

sch_gss <- paste(as.character(use_dat$school_id), as.character(use_dat$gss_code))

final_dat <- data.frame("ID" = sch_gss, "year" = as.numeric(use_dat$year), "y" = as.numeric(use_dat$ft_tot_all_races_v))
example_index <- sample(unique(final_dat$ID), 50, replace = FALSE)
final_dat_ex <- final_dat[final_dat$ID %in% example_index, ]
```


```{r}
ggplot(data = final_dat_ex, aes(x = year, y = y, group = ID, color = ID)) +
  geom_line() +
  labs(x = "year", y = "y", title = "Time Series Plot by School ID and GSS Code") +
  theme_minimal() + 
  scale_x_continuous(breaks = 1972:2021) + 
  theme(legend.position = "none")


```



